\Large Exponential Random Graph Models with Big Networks: Maximum Pseudolikelihood Estimation and the\\Parametric Bootstrap

### 2017-02-25 TeX/paper.tex
chapter bootstrapped MPLE

 Coming...

@@

\begin{figure}[h]
\begin{center}
\includegraphics[scale=.5]{RMSE_paper}
\caption{The log of the ratio of the RMSE for the MCMCMLE to the MPLE for different samplesizes and two different networks, Faux Mesa High and Faux Magnolia High}
\label{rmse}
\end{center}
\end{figure}

\noindent The plots support the fact that larger networks require a larger samplesize of simulated networks in order for the MCMCMLE to outperform the MPLE. While the fairly small Faux Mesa High network only requires a samplesize of about 50 networks the larger Faux Magnolia High network already requires a samplesize of at least 1500 networks for the MCMCMLE to surpass the MPLE. These results propose especially for very large networks (e.g. social media data) the question, of how large the samplesize has to be set in order to justify the approximately exact, but comutationally expensive MCMCMLE method. The identification of the required samplesize is not only becoming more difficult as the size of the network increases, but an increase of the samplesize will also extends the mere computation time.

\section*{Bootstrapped MPLE}
\begin{itemize}

>>>>

 Coming soon...

@@

\begin{figure}[h]
\begin{center}
\includegraphics[scale=.5]{RMSE_paper}
\caption{The log of the ratio of the RMSE for the MCMCMLE to the MPLE for different sample sizes and two different networks, Faux Mesa High and Faux Magnolia High}
\label{rmse}
\end{center}
\end{figure}

\noindent The plots support the fact that larger networks require a larger sample size of simulated networks in order for the MCMCMLE to outperform the MPLE. While the fairly small Faux Mesa High network only requires a sample size of about 50 networks the larger Faux Magnolia High network already requires a sample size of at least 1500 networks for the MCMCMLE to surpass the MPLE. These results propose especially for very large networks (e.g. social media data) the question, of how large the sample size has to be set in order to justify the approximately exact, but computationally expensive MCMCMLE method. The identification of the required sample size is not only becoming more difficult as the size of the network increases, but an increase of the sample size will also extends the mere computation time.

\section*{Bootstrapped MPLE}
As discussed in the previous section, the MPLE converges to the MLE as the size of the network increases. Moreover, it is able to outperform the MCMCMLE if sample size is not chosen large enough. The main reason, why the MCMCLE is still widely preferred is, that in contrast to the MPLE, it does not underestimate the variance of its estimates. By the definition of the ERGM it is obvious that this model is an exponential family of distributions where $\theta$ is the natural parameter and $\Gamma (Y)$ takes the part of the sufficient statistic. This allows us to work in the framework of exponential family distributions and as a consequence, leads to the conclusion that the sampling distribution of the MLE is multivariate normal with mean vector equal to the MLEs and a covariance matrix equal to the inverse of the negative Hessian matrix $[-H]^{-1}$ of the likelihood function at the MLE. The problem with the MPLE is that the inverse of the negative hessian matrix calculated by the pseudolikelihood function will underestimate the variance of the MPLE \cite{Duijnetall2009}, resulting in an underestimate of the width of the confidence intervals. Duijn et all (cite) show that constructing 95\% MPLE confidence intervals can result in intervals that only comprise the true value in less than 75\% instead of the demanded 95\%. In this paper, we are going to refer to the MPLE confidence intervals as \textit{logistic regression confidence intervals} simply because the MPLE is calculated using logistic regression methods that also use $[-H]^{-1}$ as an estimate for the covariance matrix. \\
Since the MPLE has the advantage of being approximately exact and computationally inexpensive, but has the disadvantage of underestimating corresponding confidence intervals, we are going to apply a technique, referred to as \textit{bootstrapped MPLE}, that was first introduced by \cite{Desmarais.2012} and provides a consistent estimate of MPLE confidence intervals. Desmarais and Cranmer \cite{Desmarais.2012} also show that this bootstrapped MPLE is a multivariate \textit{M}-estimator, a special class of robust estimators, proving that this algorithm consistently estimates the confidence intervals of the MPLE.
In order to calculate the bootstrapped MPLE, a sample of $n$ networks are being drawn with replacement from the distribution of $Y$. Then, we evaluate the MPLE on every sampled network and take the $2.5$th and $97.5$th percentile of the MPLE estimates to obtain 95\% bootstrap confidence intervals. Obviously, we obtain bootstrap confidence intervals with a different significance level $\alpha$ by simply picking the corresponding percentiles.

\begin{itemize}
